### 2025-03-06 template.qmd
done here

---
title: Recent Advances in Underwater Basket Weaving Under the Extreme Pressure of the Mariana Trench
title-running: Underwater Basket Weaving Under Extreme Pressure
format:
  quarto-ecml-pdf:
    keep-tex: true
  quarto-ecml-html: default
blind-review: false
author:
  - name: André Lauren Benjamin
    affiliations:
      - name: Fictional Southern University
        address: Savannah GA 31404
        country: USA
        url: \{a.l.benjamin,a.a.patton\}\@fsu.fake
        id: south
    url: https://example.org/
  - name: Calvin Cordozar Broadus Jr.
    corresponding: true 
    affiliations:
      - name: Fictional West Coast University
        address: Long Beach CA 90840
        country: USA
        url: ccb@fwcu.fake
      - name: Secondary European Affiliation
        address: Tiergartenstr. 17
        city: Heidelberg
        country: Germany
        postal-code: 69121 
        url: lncs@springer.com
  - name: Antwan André Patton
    orcid: '0000-1111-2222-3333'
    affiliations:
      - ref: south 
author-running: A.L. Benjamin et al.
abstract: |
  The abstract should briefly summarize the contents of the paper in 150--250 words.
keywords: [First Keyword, Second Keyword, Another Keyword.]
---

## Introduction {#sec-intro}

### Main Contributions

Please note that the first paragraph of a section or subsection is not indented. The first paragraph that follows a table, figure, equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

## Related Work

### Basket Weaving

#### Underwater Basket Weaving

Only two levels of headings should be numbered. Lower level headings remain unnumbered; they are formatted as run-in headings.

##### Underwater Basket Weaving Under Difficult Circumstances

The contribution should contain no more than four levels of headings. @tbl-headings gives a summary of all heading levels.

::: {#tbl-headings tbl-pos='t'}
```{=latex}
\begin{tabular}{lll}
\toprule
Heading level &  Example & Font size and style\\
\midrule
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\bottomrule
\end{tabular}
```

Table captions should be placed above the tables.
:::

<!-- | Heading level | Example | Font size and style |
|------|------|------|
| Title (centered) | {\Large\bfseries Lecture Notes} | 14 point, bold |
| 1st-level heading | {\large\bfseries 1 Introduction} | 12 point, bold |
| 2nd-level heading | {\bfseries 2.1 Printing Area} | 10 point, bold |
| 3rd-level heading | {\bfseries Run-in Heading in Bold.} Text follows | 10 point, bold |
| 4th-level heading | {\itshape Lowest Level Heading.} Text follows | 10 point, italic |

: Table captions should be placed above the
tables. {#tbl-headings} -->

## Recent Advances from the Mariana Trench

Displayed equations are centered and set on a separate line.

$$
x + y = z
$$ {#eq-linear}
Please try to avoid rasterized images for line-art diagrams and schemas. Whenever possible, use vector graphics instead (see
@fig-duck).

![A figure caption is always placed below the illustration. Please note that short captions are centered, while long ones are justified by the macro package automatically.](example-image-duck){#fig-duck width=100% fig-pos='t'}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
<!-- %
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
% -->
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}

<!-- Using Quarto theorem environments does not work, see here: https://github.com/pat-alt/quarto-ecml/issues/1 -->
<!-- ::: {#thm-sample}
This is a sample theorem. The run-in heading is set in bold, while the following text appears in italics. Definitions, lemmas, propositions, and corollaries are styled the same way.
:::

::: {.proof}
Proofs, examples, and remarks have the initial word in italics, while the following text appears in normal font.
::: -->

For citations of references, we prefer the use of square brackets and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides a sample reference list with entries for journal articles [@ref_article1], an LNCS chapter [@ref_lncs1], a book [@ref_book1], proceedings without editors [@ref_proc1], and a homepage [@ref_url1]. Multiple citations are grouped [@ref_article1;@ref_lncs1;@ref_book1], [@ref_article1;@ref_book1;@ref_proc1;@ref_url1].

## Experiments

### Experimental Setup

### Experimental Results

## Discussion

## Conclusion

Of course, authors have complete freedom on how they choose to structure their paper. Section headers from Introduction up to and including Conclusions are completely up to the discretion of the authors; use whichever structure you see fit. Title, Abstract, the credits environment, and References, however, are mandatory.

\begin{credits}
\subsubsection{\ackname} A bold run-in heading in small font size at the end of the paper is
used for general acknowledgments, for example: This study was funded
by X (grant number Y).

\subsubsection{\discintname}
It is now necessary to declare any competing interests or to specifically
state that the authors have no competing interests. Please place the
statement with a bold run-in heading in small font size beneath the
(optional) acknowledgments,
for example: The authors have no competing interests to declare that are
relevant to the content of this article. Or: Author A has received research
grants from Company W. Author B has received a speaker honorarium from
Company X and owns stock in Company Y. Author C is a member of committee Z.
\end{credits}